Topic modeling on EU and the USA research projects

### 2017-06-24 thesis_report/draft_thesis.tex
Introduction

\begin{abstract}
The purpose of this work is to better identify, quantify and understand themes
in research and innovation over the last 22 years, and to provide valuable
knowledge about the research trends nowadays. A way to approach the issue is
through the projects supported by funding agencies. In particular, public data
are available for the projects funded under the European Union Framework
Programmes, and for project funder by the NSF in the United States. The data
include project descriptions, so it is possible to investigate, based on these
textual data, the research topics that were actually funded over the last two
decades. We first introduce the concept of Latent Dirichlet allocation and its
application in topic modelling in projects funded in the two regions from 1994
to 2017. We will then proceed to show the differences between the two regions in
terms of topic similarity and topic evolution over time. We will discuss both
policy implications, what has been funded over the years, as well as technical
ones, how can one handle such datasets and extract information from them.
\end{abstract}

\section{Introduction}
This is where you tell people why they should bother reading your article.

@@

Moreover, the information about what has been funded over the last years in
these two regions show that the scope of scientific research has become more
interdisciplinary and international as well. Even though we started analysing
the abstracts of the projects quantitatively, we also tried to discuss the
results and explain furtherthe evolution of the topics in research and potential
trends.

To sum up, discovering the topics in scientific research is a first step to
understand the knowledge that the research offers to the community in these two
regions, visualise the content and discover promising trends. We hore our work
could help people realise the potential of scientific research and inspire them
to identify meaningful topics to focus on.

>>>>

\begin{abstract}
The purpose of this work is to better identify, quantify and understand themes
in research and innovation over the last 22 years, and to provide valuable
knowledge about the research trends nowadays. A way to approach the issue is
through the projects supported by funding agencies. In particular, public data
are available for the projects funded under the European Union Framework
Programmes, and for project funded by the NSF in the United States. The data
include project descriptions, so it is possible to investigate, based on these
textual data, the research topics that were actually funded over the last two
decades. We first introduce the concept of Latent Dirichlet allocation and its
application in topic modeling in projects funded in the two regions from 1994
to 2017. We will then proceed to show the differences between the two regions in
terms of topic similarity and topic evolution over time. We will discuss both
policy implications, what has been funded over the years, as well as technical
ones, how can one handle such datasets and extract information from them.
\end{abstract}

\section{Introduction}
Research is the foundation on which knowledge and new technologies emerge and
support our welfare. Scientific research brings together observation, knowledge
and data to solve problems, invent solutions and develop new products. In our
globalized world, the shared commitment for excellence in science has a major
impact to our well being and it's important to support research and development.
Over the last years, many projects from different domains are getting funded for
research and innovation.

Nowadays, there are many funding opportunities for researchers. For instance,
European Union countries are encouraged to invest 3\% of their GDP in research
and development by 2020. EU's latest research programme, Horizon 2020, allows
researchers to identify new opportunities and directions in almost any field of
research. The programme's goals are to strengthen the EU's position in science,
support industrial innovation and address social concerns, such as climate
change, green transportation and energy.

Another funding agency that actively promotes the progress of science is the
National Science Foundation in the United States of America. NSF supports
research in all the non medical fields of science \& engineering and aims to
advance the national health and prosperity. It finds approximately 24\% of the
federally supported research conducted by the USA's colleges and universities.

With our work, we aim to investigate the evolution of scientific research in the
EU and the US. In other words, we attempt to identify which subjects are more
likely to get a funding, based on the previous focus the two regions in research
fields. In order to answer these questions, we tried to gain some insights from
the projects that got funded the last two decades under the EU Framework
Programmes and the NSF in USA.

When researchers apply for a funding, one of the first steps they do is to
identify an interesting subject of the many possible topics of scientific
investigation. The topic(s) of a project is always described in the project's
abstract/description. Obtaining these abstracts, one can possibly investigate
the topics of the projects and find the subjects that are being addressed and
gain some information regarding the popularity of the topics among others.

In this work, we discover topics and their evolution from 1994 to 2017, using
the projects' abstracts that got funded this period. Using a topic model, we
extract a set of topics and investigate the changes in the content per
year/framework programme. These topics are used to gain insight into some of the
research that occurs the last two decades. Our goal is to share these insights
with researchers and encourage them to develop better strategies to establish
themselves as more effective researchers globally, by choosing a subject that is
more likely to get a funding.

In section 2, we briefly talk about the papers that mostly influenced our study.
Section 3 introduces some basic concepts in topic modeling and we describe the
model that we used for topic extraction, Latent Dirichlet Allocation, which is a
statistical generative model for documents. In section 4 we provide information
about the dataset structure and the methodology that we followed in our
analysis. In section 5 we interpret the results of our analysis, discuss the
evolution of the topics over time and the differences between the two regions'
topics based on their similarity. Finally, section 6 summarizes our work and
section 7 suggests some future work.

@@

Moreover, the information about what has been funded over the last years in
these two regions show that the scope of scientific research has become more
interdisciplinary and international as well. Even though we started analyzing
the abstracts of the projects quantitatively, we also tried to discuss the
results and explain further evolution of the topics in research and potential
trends.

To sum up, discovering the topics in scientific research is a first step to
understand the knowledge that the research offers to the community in these two
regions, visualize the content and discover promising trends. We hope our work
could help people realize the potential of scientific research and inspire them
to identify meaningful topics to focus on.
